---
title: "a4-task2-mukta-kelkar"
author: "Mukta Kelkar"
date: "11/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(lubridate)
library(kableExtra)
library(effsize)
library(ggbeeswarm)
```

# Introduction

[PLACEHOLDER TEXT]

# Summary of Data and Analyses

[PLACEHOLDER TEXT]

```{r read in data and do initial wrangling, include=FALSE}
# Read in data
bonanza_hares <- read.csv(here("data", "bonanza_hares.csv"))

# Use filter function to chose only juvenile hares
juvenile_hares <- bonanza_hares %>% 
  filter(age == "j")

# Sort data by year
hares_by_year <- juvenile_hares %>% 
  mutate(mdy(date)) %>% 
  rename(date_form = "mdy(date)") %>% 
  mutate(year = year(date_form))
```

# Exploratory Findings

## Annual Juvenile Hare Trap Counts

```{r summarize data for juvenile hare counts, include=FALSE}
# Count the total number of juvenile (age == "j") hare trappings during each year of the study
juveniles_per_year <- hares_by_year %>% 
  count(age, year)

# Assess summary statistics
summary(juveniles_per_year)

# Find summary statistics for use in report out
juvenile_summary <- juveniles_per_year %>% 
  summarise(
    mean_count = mean(n),
    median_count = median(n),
    min_count = min(n),
    max_count = max(n)
  )
```

```{r create column graph, echo = FALSE, message = FALSE, warning = FALSE, fig.align = "center"}
# Create a final visualization of counts by year
ggplot(data = juveniles_per_year, aes(x = year, y = n)) +
  geom_col() +
  labs(y = "Number of Juveniles",
       x = "Year",
       title = "Number of juvenile bonanza hares per year",
       caption = "Figure 1: Juvenile Hares caught in traps along Bonanza Creek between 1999 and 2012") +
  theme_minimal()
```


Overall, it looks like the number of juvenile hares caught in traps has decreased with time. The median count is `r juvenile_summary$median_count` and the mean count is `r juvenile_summary$mean_count`. The minimum number of hares is `r juvenile_summary$min_count` in the year 2010 and the maximum number of hares is `r juvenile_summary$max_count` in the year 1999. One consideration to note when analyzing the data is that the count is impacted by effort (in other words, the number of days and number of traps in the study each year). A possible way to address this moving forward would be to divide the number of hares caught per year by the number of traps deployed per year. Then, instead of looking at number of hares per year, we can look at number of hares per trap per year. 


## Juvenile Hare Weights

```{r wrangle data, include=FALSE, warning=FALSE, message = FALSE}
# Create dataset that includes only variables age, sex, and site(grid)
j_hare_weights <- juvenile_hares %>% 
  select("grid", "sex", "age", "weight") %>% 
  group_by(grid) %>% 
  mutate(site_name = case_when(      #create new column with more ckear site names
    grid == "bonrip" ~ "Bonanza Riparian",
  grid == "bonmat" ~ "Bonanza Mature",
  grid == "bonbs" ~ "Bonanza Black Spruce"
  ))

# View summary of data
summary(j_hare_weights)

```

```{r visualize data, echo = FALSE, message = FALSE, warning = FALSE, fig.align = "center"}
# Create a final visualization
ggplot(data = j_hare_weights,
       aes(x = sex, y = weight)) +
  geom_beeswarm(aes(color = sex),  #add a beeswarm plot in the first layer of the graph
                size = 2,
                alpha = 0.8) +
  facet_wrap(~site_name) +    #wrap the data by study site
  theme_minimal() +
  geom_boxplot(data = j_hare_weights, aes(x = sex, y = weight), #add a boxplot in the second layer of the graph
               size = 1,
               alpha = 0.1,
               color = "gray30") +
  theme(legend.position = "none") +  #remove the legend from the graph
  labs(x = "Sex",
       y = "Weight",
       title = "Variation in juvenile hare weights by sex and study site",
       caption = "Figure 2: Weight of juvenile hares caught in traps at three different sites along Bonanza Creek between 1999 and 2012.")
```


Prior to running any statistical analyses, there is minimal variation between the mean weight between sites. It looks like male juveniles may weigh more than female juvenile hares, which I examine statistically in the following sections. 


## Juvenile weight comparison by sex

```{r, include = FALSE}
# Take out NAs
j_hare_weights_clean <- j_hare_weights %>%
  filter(!is.na(weight)) %>%
  filter(!is.na(sex))

#Create vectors for statistical tests
hare_f <- j_hare_weights_clean %>%   #create vector for weights of juvenile female hares
  filter(!is.na(sex)) %>%
  filter(sex == "f") %>%
  pull(weight)
hare_m <- j_hare_weights_clean %>%   #create vector for weights of juvenile male hares
  filter(!is.na(sex)) %>%
  filter(sex == "m") %>%
  pull(weight)

# Create data frame of summary statistics
hare_weight_table <- j_hare_weights_clean %>% 
  group_by(sex) %>% 
  summarize(
    mean__weight = round(mean(weight, na.rm = TRUE),2),
    sd_hares = round(sd(weight, na.rm = TRUE),2),
    n = n()
  )

# Make a histogram that shows the distribution of weights for both female and male juvenile hares
ggplot(data = j_hare_weights_clean, aes(x = weight)) +  #
  geom_histogram(bins = 20) +
  facet_wrap(~sex)

# Make a qq plot to check for normality
ggplot(data = j_hare_weights_clean, aes(sample = weight)) + 
  geom_qq() +
  facet_wrap(~sex) #weights of both sexes are normally distributed

# Run a t-test
hare_ttest <- t.test(hare_f, hare_m)

# Get the effect size: 
hare_effsize <- cohen.d(hare_f, hare_m)
```

```{r data table, , echo = FALSE, message = FALSE, warning = FALSE, table.align = "center"}
# Format data frame as a table using kalbeExtra
hare_weight_table %>% 
  kable(col.names = c("Sex", "Mean Weight", "Standard Deviation", "Sample Size")) %>% 
  kable_styling(bootstrap_options = "striped", 
                full_width = FALSE)
```

Overall, male juvenile hares weigh more than female juvenile hares. The actual difference in means is `r hare_weight_table[1,1] - hare_weight_table[1,2]` and the percent diference is [PLACEHOLDER]. The difference in means is significant (F statistic = `r round(hare_ttest$statistic,2)`, p value = `r round(hare_ttest$p.value,2)`, p < 0.05) but the effect size is small (Cohen's *d* = `r round(hare_effsize$estimate,2)`).


## Relationship between juvenile weight & hind foot length

```{r}
# Wrangle data to include only hind foot length and weight of juvenile hares. Then filer out the NA values for both hind food length and weight.
j_weight_length <- juvenile_hares %>% 
  select(hindft, weight) %>% 
  filter(!is.na(weight)) %>%
  filter(!is.na(hindft))

# Create a line graph that shows the relationship between hind food length and weight
ggplot(data = j_weight_length, aes(x = hindft, y = weight)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm",
              color = "red",
              size = 0.5,
              fill = "gray10",
              alpha = 0.5) +
  labs(x = "Hind foot length (mm)",
       y = "Weight",
       title = "Relationship between weight and hind foot length in juvenille hares",
       subtitle = "Bonanza Creek Experimental Forest: 1999 - Present",
       caption = "Figure 4: ahhhhhhhhh") +
  theme_bw() +
  ggpubr::stat_regline_equation(label.x = 80, label.y = 1700)

# Linear model, stored as j_hare_lm
j_hare_lm <- lm(hindft ~ weight, data = j_weight_length)

# Return the complete overview:
summary(j_hare_lm)
```


# Final Summary 

# Citations