---
title: "a4-task2-mukta-kelkar"
author: "Mukta Kelkar"
date: "11/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(lubridate)
library(kableExtra)
library(effsize)
library(ggbeeswarm)
```

# Introduction

[PLACEHOLDER TEXT]

# Summary of Data and Analyses

[PLACEHOLDER TEXT]

```{r read in data and do initial wrangling, include=FALSE}
# Read in data
bonanza_hares <- read.csv(here("data", "bonanza_hares.csv"))

# Use filter function to chose only juvenile hares
juvenile_hares <- bonanza_hares %>% 
  filter(age == "j")

# Sort data by year
hares_by_year <- juvenile_hares %>% 
  mutate(mdy(date)) %>% 
  rename(date_form = "mdy(date)") %>% 
  mutate(year = year(date_form))
```

# Exploratory Findings

## Annual Juvenile Hare Trap Counts

```{r summarize data for juvenile hare counts, include=FALSE}
# Count the total number of juvenile (age == "j") hare trappings during each year of the study
juveniles_per_year <- hares_by_year %>% 
  count(age, year)

# Assess summary statistics
summary(juveniles_per_year)

# Find summary statistics for use in report out
juvenile_summary <- juveniles_per_year %>% 
  summarise(
    mean_count = mean(n),
    median_count = median(n),
    min_count = min(n),
    max_count = max(n)
  )
```

```{r create column graph, echo = FALSE, message = FALSE, warning = FALSE, fig.align = "center"}
# Create a final visualization of counts by year
ggplot(data = juveniles_per_year, aes(x = year, y = n)) +
  geom_col() +
  labs(y = "Number of Juveniles",
       x = "Year",
       title = "Number of juvenile bonanza hares per year",
       caption = "Figure 1: Juvenile Hares caught in traps along Bonanza Creek between 1999 and 2012") +
  theme_minimal()
```


Overall, it looks like the number of juvenile hares caught in traps has decreased with time. The median count is `r juvenile_summary$median_count` and the mean count is `r juvenile_summary$mean_count`. The minimum number of hares is `r juvenile_summary$min_count` in the year 2010 and the maximum number of hares is `r juvenile_summary$max_count` in the year 1999. One consideration to note when analyzing the data is that the count is impacted by effort (in other words, the number of days and number of traps in the study each year). A possible way to address this moving forward would be to divide the number of hares caught per year by the number of traps deployed per year. Then, instead of looking at number of hares per year, we can look at number of hares per trap per year. 


## Juvenile Hare Weights

```{r, include=FALSE, warning=FALSE, message = FALSE}
# Create dataset that includes only variables age, sex, and site(grid)
j_hare_weights <- juvenile_hares %>% 
  select("grid", "sex", "age", "weight") %>% 
  group_by(grid) %>% 
  mutate(site_name = case_when(
    grid == "bonrip" ~ "Bonanza Riparian",
  grid == "bonmat" ~ "Bonanza Mature",
  grid == "bonbs" ~ "Bonanza Black Spruce"
  ))

# View summary of data
summary(j_hare_weights)

# Create a boxplot
ggplot(data = j_hare_weights,
       aes(x = sex, y = weight)) +
  geom_beeswarm(aes(color = sex),
                size = 2,
                alpha = 0.8) +
  facet_wrap(~site_name) +
  theme_minimal() +
  geom_boxplot(data = j_hare_weights, aes(x = sex, y = weight),
               size = 1,
               alpha = 0.1,
               color = "gray30") +
  theme(legend.position = "none") +
  labs(x = "Sex",
       y = "Weight",
       title = "Variation in juvenile hare weights by sex and study site",
       caption = "Figure 2: Weight of juvenile hares caught in traps at three different sites along Bonanza Creek between 1999 and 2012.")


```
Prior to running any statistical analyses, there is minimal variation between the mean weight and distribution of weight between sites. It looks like male juveniles may weigh more than female juvenile hares, which I examine statistically in the following sections. 


## Juvenile weight comparison by sex

```{r, include = FALSE}
# Take out NAs
j_hare_weights_clean <- j_hare_weights %>%
  filter(!is.na(weight)) %>%
  filter(!is.na(sex))

#Create vectors for statistical tests
hare_f <- j_hare_weights_clean %>%
  filter(!is.na(sex)) %>%
  filter(sex == "f") %>%
  pull(weight)
hare_m <- j_hare_weights_clean %>%
  filter(!is.na(sex)) %>%
  filter(sex == "m") %>%
  pull(weight)

# Create table of summary statistics
hare_weight_table <- j_hare_weights_clean %>% 
  group_by(sex) %>% 
  summarize(
    mean__weight = round(mean(weight, na.rm = TRUE),2),
    sd_hares = round(sd(weight, na.rm = TRUE),2),
    n = n()
  )

# Format table useing kalbeExtra
hare_weight_table %>% 
  kable(col.names = c("Sex", "Mean Weight", "Standard deviation", "Sample size")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

#Check for normality

# Run a t-test
t.test(hare_f, hare_m)

# Get the effect size: 
hare_effsize <- cohen.d(hare_f, hare_m)
```


## Relationship between juvenile weight & hind foot length

```{r}
# Wrangle data to include only hind foot length and weight of juvenile hares. Then filer out the NA values for both hind food length and weight.
j_weight_length <- juvenile_hares %>% 
  select(hindft, weight) %>% 
  filter(!is.na(weight)) %>%
  filter(!is.na(hindft))

# Create a line graph that shows the relationship between hind food length and weight
ggplot(data = j_weight_length, aes(x = hindft, y = weight)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm",
              color = "red",
              size = 0.5,
              fill = "gray10",
              alpha = 0.5) +
  labs(x = "Hind foot length (mm)",
       y = "Weight",
       title = "Relationship between weight and hind foot length in juvenille hares",
       subtitle = "Bonanza Creek Experimental Forest: 1999 - Present",
       caption = "Figure 4: ahhhhhhhhh") +
  theme_bw() +
  ggpubr::stat_regline_equation(label.x = 80, label.y = 1700)

# Linear model, stored as j_hare_lm
j_hare_lm <- lm(hindft ~ weight, data = j_weight_length)

# Return the complete overview:
summary(j_hare_lm)
```


# Final Summary 

# Citations